Review of Assigned Readings

### 2018-01-05 paper/paper.tex
final reading of paper.tex and minor grammar/spelling changes

\textcite[p.~388]{malkiel1970efficient} introduced the Efficient Market Hypothesis (EMH), where they claim that under certain market conditions actual prices include all information. This hypothesis is broadly accepted in the financial world. The digitalisation leads to an increased network effect, where information can be shared over the globe. Based on \citeauthor{mao2015quantifying} (\citeyear[][p.~3]{mao2015quantifying}, as cited in \cite[][pp.~175--195]{shiller2015irrational}; \cite[][pp.~279]{kahneman2013prospect}) the EMH fails to address the behavioural and emotional role of investors. The large fluctuation of the Bitcoin prices in the last year (mostly increasing) and the fact that Twitter has become a very popular social network, where people interact, led to different research in the field of sentimental analysis with the focus on Bitcoin and Twitter. \textcite[p.~18]{mao2015quantifying}, claim that sentiment analysis using Twitter tweets which contain the buzzword "bullishness" can indeed be used as a sentiment indicator for stock prices. As the modern economy becomes more digitised, crypotocurrencies, as a financial asset, draw more attention. This paper seeks to shed light on whether the tweets can be used to predict Bitcoin prices.

@@

\subsection{Methodology}
\label{sec:Methodology}
To answer the above mentioned research question, a scientific approach has been applied. The paper consists out of three parts. First, the relevance of the topic is pointed out by the introduction , the research question, methodology and the scope (see chapter \ref{sec:intro}). Secondly, the related theory has been analysed to provide the reader with the necessary background information by briefly introducing both Bitcoin and Twitter in context of price prediction and behavioural finance. The covered research is based on forward research in regard to the stock market and cryptocurrencies with a special focus on sentiment analysis and Bitcoin \ref{sec:Background}. Further on,  Vector Autoregression (VAR) approach is introduced and its constrains in context of the paper are being outlined (see chapter \ref{sec:EconometricModellingandResults}). The third part contains the discussion of the results. The computation is based on a data set, which was gathered by Python scripts for the time period December 20th - 27th. Since the results are based on the observation of a short period, they have to be interpreted carefully. The data set consists of two data sources with the following characteristics: (1) Twitter tweets which contain the buzzword "Bitcoin" and (2) those containing historical Bitcoin prices. The two sources have been aggregated by using Python. The output of the aggregation was further processed with STATA. The third part of the paper is discussed in the chapter \ref{sec:EconometricModellingandResults} followed by a brief conclusion in chapter \ref{sec:Conclustion}.

\subsection{Scope}
\label{sec:Scope}
The paper's focus lies on the scientific discussion of empirical test results. The research question has consciously been narrowed,in order to build up on a existing methodology for sentiment analysis in the stock market while simultaneously extending the research scope in the area of behavioural finance and sentiment analysis for cryptocurrencies. The short time frame of our project further limited the period of data collection which has a direct impact on the reliability of our results.\\
Further on, this paper should act as a blueprint for further research in the area of Twitter sentiment analysis for cryptocurrencies. Therefore, the accompanying project documentation can be used as a guideline for further research.

@@

In previous research \textcite{madan2015automated} applied \textit{machine learning} algorithms to predict Bitcoin prices. By using the \textit{bayesian regression}, \textcite{shah2014bayesian} achieved a return on investment of 89 percent over an investment period of 50 days. However, research in this area is still limited and most of them do not take into consideration the individuals' sentiments (\cite[p.~1]{colianni2015algorithmic}).\\

One core assumption of behavioural finance is that investors’ interests influence their behaviour and therefore stock prices (\cite[p.~3]{mao2015quantifying}). A widely used approach to measure investors’ interests is the sentiment analysis of Twitter data, which will be covered in the following section.\\

@@

Twitter was founded in 2006 and gained rapidly worldwide popularity and reached 330 million active users in the third quarter of 2017  (\cite[]{twitterinc2018}; \cite[]{Statista2018b}). As a micro-blogging platform, Twitter suits the purpose of analysing investors’ interests due to its special characteristics. In comparison to other social networks, Twitter limits its posts (tweets) to relatively short messages of 140 characters. Tweets can contain observations, thoughts, links to interesting content, websites, uploaded pictures or videos. Special conventions such as hashtags “\#” or mentions “@” are used to reference searchable content or user profiles.(\cite{Schmidt2018socialmedia})\\

@@

\subsection{Data Collection}
For our data, we collected two separate time series datasets. One showing the Twitter activity regarding Bitcoin, with the second one showing the historical Bitcoin price development. Both datasets were collected hourly for the period from the 19.12.2017 starting at 20:00:00 UTC and ending on the 28.12.2017 at 10:00:00 UTC. We chose the hourly basis to conduct statistical analysis with a limited amount of observations as demanded by the feasibility of such a complex project. For the Twitter data we aggregated by each hour all tweets mentioning "Bitcoin". We used prices at the last second of every hour to represent the hourly Bitcoin price.\\
The following figure displays the Twitter activity with regard to Bitcoin for the observed time period. The first plot shows the hourly change of the aggregate number of tweets containing "Bitcoin". In the second and third plots we respectively did a first difference and a logarithmic first difference transformation of our main variable, in case it is needed in the later analysis. The graphs show overall a cyclical movement around 1200 tweets and a single outburst befalling on 22.12.2017 between 15:00:01 and 21:00:00, with highest value being 4622 tweets in one hour.\\

@@

Next we display the time series data on Bitcoin price development for the observed time period. The first plot shows the hourly changes of Bitcoin prices. The second and third plots are once again first difference transormations of the main variable for later analysis. In this graph we observe a notably volatile movement with no spesific direction. This contradicts the general rising trend of Bitcoin prices of the past months, but the problem may indeed arise if we happen to select a short period of one week when Bitcoin price movement behaves less bullishly as usual, in other words, the lack of articulated trend may be attributed to \textit{selection bias}.\\

@@

%************************** Chapter 3 - Econometric Modelling and Results **************************
\section{Econometric Modelling and Results}
\label{sec:EconometricModellingandResults}
As stated in the beginning, our purpose is to examine the underlying relationship between the number of tweets containing "Bitcoin" and Bitcoin price fluctuations. From our data collection, we are left with two time series datasets representing two variables. A broadly used statistical method to simultaneously analyse multiple time series is the Vector Autoregression (VAR) approach. In this method the endogenous variables are determined both by their own historical values and by the historical values of the other endogenous variables (\cite[pp.~4--5]{luetkepohl2007new}).\\
To generate the following econometric results we used the statistical software STATA.

\subsection{Preparing for VAR: Stationarity Test}
Before estimating a VAR model, the time series data's stationarity must be confirmed. Namely, we need to demonstrate that the means and variances are constant over time, and that the dataset does not show any trending behavior. Non-stationary data can lead to an inaccurate estimation with VAR model. To conduct stationarity test, we use the Augmented Dickey-Fuller (ADF) test.\\

We shall test the stationarity for the "crude" Twitter data (variable name: nr\_of\_tweets), and since the variable may not pass the stationarity test, we also check for the stationarity of the first difference Twitter data (df\_nr\_of\_tweets) and the logarithmic first difference one (log\_df\_nr\_of\_tweets). The results of the augmented Dickey-Fuller (ADF) test are listed in the following table:\\

@@

From the results we can see that only the first variable (i.e. the untransformed tweet count) is non-stationary, since the absolute value of the Test Statistic (3.152) is smaller than the absolute value of the 1\% critical value (3.475), the null hypothesis of non-stationarity cannot be rejected as a consequence. The second variable is stationary (14.167 $>$ 3.475) and from here onwards we shall proceed the analysis with only the first difference tweet count, whose test statistic falls furthest into the rejection zone.\\

Out of the similar concern stated above, we shall check for stationarity for the second variable, the Bitcoin prices (variable name: bpi\_closing\_price). As before, we also include the first difference Bitcoin prices (variable name: df\_bpi\_closing\_price) and the logarithmic first difference ones (variable name: log\_df\_bpi\_closing\_price) transformations. The ADF test for these variables gives the following results:\\

@@

\begin{table}
\centering
	\begin{tabular}{lcc} \hline
	 & (T\textsubscript{t}) & (B\textsubscript{t}) \\
	VARIABLES & df\_nr\_of\_tweets & df\_bpi\_closing\_price \\ \hline
	 &  &  \\
	L.df\_nr\_of\_tweets (T\textsubscript{t-1}) & 0.00999 (a\textsubscript{11}) & 0.00463 (a\textsubscript{21}) \\
	 & (0.0693) & (0.0104) \\
	L.df\_bpi\_closing\_price (B\textsubscript{t-1}) & -0.675 (a\textsubscript{12}) & -0.125* (a\textsubscript{22}) \\
	 & (0.469) & (0.0707) \\
	Constant  & -88.34 (c\textsubscript{1}) & -18.48 (c\textsubscript{2}) \\
	 & (177.9) & (26.79) \\
	 &  &  \\
	 Observations & 205 & 205 \\ \hline
	\multicolumn{3}{c}{ Standard errors in parentheses} \\
	\multicolumn{3}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
	\end{tabular}
\caption{Var Regression}
\end{table}

>>>>

\textcite[p.~388]{malkiel1970efficient} introduced the Efficient Market Hypothesis (EMH), where they claim that under certain market conditions actual prices include all information. This hypothesis is broadly accepted in the financial world. The digitalisation leads to an increased network effect, where information can be shared over the globe. Based on \citeauthor{mao2015quantifying} (\citeyear[][p.~3]{mao2015quantifying}, as cited in \cite[][pp.~175--195]{shiller2015irrational}; \cite[][pp.~279]{kahneman2013prospect}) the EMH fails to address the behavioral and emotional role of investors. The large fluctuation of the Bitcoin prices in the last year (mostly increasing) and the fact that Twitter has become a very popular social network, where people interact, led to different research in the field of sentimental analysis with the focus on Bitcoin and Twitter. \textcite[p.~18]{mao2015quantifying}, claim that sentiment analysis using Twitter tweets which contain the buzzword "bullishness" can indeed be used as a sentiment indicator for stock prices. As the modern economy becomes more digitised, cryptocurrencies, as a financial asset, draw more attention. This paper seeks to shed light on whether the tweets can be used to predict Bitcoin prices.

@@

\subsection{Methodology}
\label{sec:Methodology}
To answer the above mentioned research question, a scientific approach has been applied. The paper consists out of three parts. First, the relevance of the topic is pointed out by the introduction, the research question, methodology and the scope (see chapter \ref{sec:intro}). Secondly, the related theory has been analysed to provide the reader with the necessary background information by briefly introducing both Bitcoin and Twitter in context of price prediction and behavioral finance. The covered research is based on forward research in regard to the stock market and cryptocurrencies with a special focus on sentiment analysis and Bitcoin \ref{sec:Background}. Further on,  Vector Autoregression (VAR) approach is introduced and its constrains in context of the paper are being outlined (see chapter \ref{sec:EconometricModellingandResults}). The third part contains the discussion of the results. The computation is based on a data set, which was gathered by Python scripts for the time period December 20th - 27th. Since the results are based on the observation of a short period, they have to be interpreted carefully. The data set consists of two data sources with the following characteristics: (1) Twitter tweets which contain the buzzword "Bitcoin" and (2) those containing historical Bitcoin prices. The two sources have been aggregated by using Python. The output of the aggregation was further processed with STATA. The third part of the paper is discussed in the chapter \ref{sec:EconometricModellingandResults} followed by a brief conclusion in chapter \ref{sec:Conclustion}.

\subsection{Scope}
\label{sec:Scope}
The paper's focus lies on the scientific discussion of empirical test results. The research question has consciously been narrowed, in order to build up on a existing methodology for sentiment analysis in the stock market while simultaneously extending the research scope in the area of behavioral finance and sentiment analysis for cryptocurrencies. The short time frame of our project further limited the period of data collection which has a direct impact on the reliability of our results.\\
Further on, this paper should act as a blueprint for further research in the area of Twitter sentiment analysis for cryptocurrencies. Therefore, the accompanying project documentation can be used as a guideline for further research.

@@

In previous research, \textcite{madan2015automated} applied \textit{machine learning} algorithms to predict Bitcoin prices. By using the \textit{bayesian regression}, \textcite{shah2014bayesian} achieved a return on investment of 89 percent over an investment period of 50 days. However, research in this area is still limited and most of them do not take into consideration the individuals' sentiments (\cite[p.~1]{colianni2015algorithmic}).\\

One core assumption of behavioral finance is that investors’ interests influence their behavior and therefore stock prices (\cite[p.~3]{mao2015quantifying}). A widely used approach to measure investors’ interests is the sentiment analysis of Twitter data, which will be covered in the following section.\\

@@

Twitter was founded in 2006 and gained rapidly worldwide popularity and reached 330 million active users in the third quarter of 2017 (\cite[]{twitterinc2018}; \cite[]{Statista2018b}). As a micro-blogging platform, Twitter suits the purpose of analysing investors’ interests due to its special characteristics. In comparison to other social networks, Twitter limits its posts (tweets) to relatively short messages of 140 characters. Tweets can contain observations, thoughts, links to various content, websites, uploaded pictures or videos. Special conventions such as hashtags “\#” or mentions “@” are used to reference searchable content or user profiles (\cite{Schmidt2018socialmedia}).\\

@@

\subsection{Data Collection}
For our data, we collected two separate time series datasets. One showing the Twitter activity regarding Bitcoin, with the second one showing the historical Bitcoin price development. Both datasets were collected hourly for the period from the 19.12.2017 starting at 20:00:00 UTC and ending on the 28.12.2017 at 10:00:00 UTC. We chose the hourly basis to conduct statistical analysis with a limited amount of observations as demanded by the feasibility of such a complex project. For the Twitter data we aggregated by each hour all tweets mentioning "Bitcoin". We used prices at the last second of every hour to represent the hourly Bitcoin price.\\
The following figure displays the Twitter activity with regard to Bitcoin for the observed time period. The first plot shows the hourly change of the aggregate number of tweets containing "Bitcoin". In the second and third plots we respectively did a first difference and a logarithmic first difference transformation of our main variable, in case it is needed in the later analysis. The graphs show overall a cyclical movement around 1200 tweets per hour and a single outburst befalling on 22.12.2017 between 15:00:01 and 21:00:00, with highest value being 4622 tweets in one hour.\\

@@

Next we display the time series data on Bitcoin price development for the observed time period. The first plot shows the hourly changes of Bitcoin prices. The second and third plots are once again first difference transformations of the main variable. In this graph we observe a notably volatile movement with no specific direction. This contradicts the general rising trend of Bitcoin prices of the past months. The problem may indeed arise if we happen to select a short period of one week when Bitcoin price movement behaves less bullishly as usual, in other words, the lack of articulated trend may be attributed to \textit{selection bias}.\\

@@

%************************** Chapter 3 - Econometric Modelling and Results **************************
\section{Econometric Modelling and Results}
\label{sec:EconometricModellingandResults}
As stated in the beginning, our purpose is to examine the underlying relationship between the number of tweets containing "Bitcoin" and Bitcoin price fluctuations. From our data collection, we are left with two time series datasets representing two variables. A broadly used statistical method to simultaneously analyze multiple time series is the Vector Autoregression (VAR) approach. In this method, the endogenous variables are determined both by their own historical values and by the historical values of the other endogenous variables (\cite[pp.~4--5]{luetkepohl2007new}).\\
To generate the following econometric results we used the statistical software STATA.

\subsection{Preparing for VAR: Stationarity Test}
Before estimating a VAR model, the time series data's stationarity must be confirmed. Namely, we need to demonstrate that the means and variances are constant over time, and that the dataset does not show any trending behavior. Non-stationary data can lead to an inaccurate estimation. To conduct stationarity, we use the Augmented Dickey-Fuller (ADF) test.\\

We shall test the stationarity for the "crude" Twitter data (variable name: nr\_of\_tweets), and since the variable may not pass the stationarity test, we also check for the stationarity of the first difference (df\_nr\_of\_tweets) and the logarithmic first difference (log\_df\_nr\_of\_tweets) transformations. The results of the augmented Dickey-Fuller (ADF) test are listed in the following table:\\

@@

From the results we can see that only the first variable is non-stationary, since the absolute value of the Test Statistic (3.152) is smaller than the absolute value of the 1\% critical value (3.475), the null hypothesis of non-stationarity cannot be rejected as a consequence. The second variable is stationary (14.167 $>$ 3.475) and from here onwards we shall proceed the analysis with only the first difference tweet count.\\

Out of the similar concern stated above, we shall check for stationarity for the second variable, the Bitcoin prices (variable name: bpi\_closing\_price). As before, we also include the first difference Bitcoin prices (variable name: df\_bpi\_closing\_price) and the logarithmic first difference (variable name: log\_df\_bpi\_closing\_price) transformations. The ADF test for these variables gives the following results:\\

@@

\begin{table}[H]
\centering
	\begin{tabular}{lcc} \hline
	 & (T\textsubscript{t}) & (B\textsubscript{t}) \\
	VARIABLES & df\_nr\_of\_tweets & df\_bpi\_closing\_price \\ \hline
	 &  &  \\
	L.df\_nr\_of\_tweets (T\textsubscript{t-1}) & 0.00999 (a\textsubscript{11}) & 0.00463 (a\textsubscript{21}) \\
	 & (0.0693) & (0.0104) \\
	L.df\_bpi\_closing\_price (B\textsubscript{t-1}) & -0.675 (a\textsubscript{12}) & -0.125* (a\textsubscript{22}) \\
	 & (0.469) & (0.0707) \\
	Constant  & -88.34 (c\textsubscript{1}) & -18.48 (c\textsubscript{2}) \\
	 & (177.9) & (26.79) \\
	 &  &  \\
	 Observations & 205 & 205 \\ \hline
	\multicolumn{3}{c}{ Standard errors in parentheses} \\
	\multicolumn{3}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
	\end{tabular}
\caption{Var Regression}
\end{table}
